---
title: "carbon model"
author: "Kemi"
date: '2022-11-07'
output: html_document
 setwd("~/Documents/JOB/Liege r sessions")
---
install.packages("deSolve")
list of all the paramters and their values
```{r}
library(deSolve)
Parms <- c (
  area = 1000000, # [m2]
  migrate= 10000, #[kg C/day] amount of salmon entering the river system
  sprate = 0.05,  # [/day] spawning plus death of salmon
  ingrate = 0.02, # [/day] bear ingestion rate
  ksSalmon = 0.01, #[kg c/m2] half saturation coeff for ingestion bear
  huntrate= 0.01/365, # [/day] death by hunting ( and other losses for bear)
  ploss = 0.5,  #[-] the loss fraction of ingested salmon to carcass
  presp = 0.4,  # [-]respired fraction of assimilated mass
  decayrate = 0.001, # [/day] decay rate of carcasses
  scavrate = 0.02,  #[/day] ingestion rate of scavengers
  ksCarcass= 0.002, #[kg C/m2] half saturation coeff for ingestion scavenger
  lossrate = 1/365  #[/day] loss rate of scavengers
)
States <- c (state <-c(BEAR = 0.01, SALMON =0, SCAVENGER = 0.005, CARCASS = 0.001))#units = kgc/m2
```

## R Markdown



```{r}
RiverRun <- function(t, state, params) {with (as.list (c(state, params)),{
  #rate expressions describing sources abd sinks- all in units of [kgc/m2/day] the functions are the equations
  Ingestion= ingrate * SALMON / (SALMON + ksSalmon) * BEAR
  Migration = migrate/area
  Spawn_death = sprate * SALMON
  Hunt_death = huntrate* BEAR
  ScavIngest = scavrate * CARCASS/ (CARCASS+ ksCarcass) * SCAVENGER
  Scavloss = lossrate * SCAVENGER
  Decay = decayrate * CARCASS
  
  #Mass balance [kgC/m2/day] for all the state variables which is a result of sources- sinks
  dBEAR = Ingestion * (1-ploss)*(1-presp) - Hunt_death
  dSALMON= Migration - Ingestion - Spawn_death
  dCARCASS = Ingestion*ploss - Decay - ScavIngest + Hunt_death
  dSCAVENGER = ScavIngest * (1-ploss)*(1-presp) -Scavloss
  list(c(dBEAR, dSALMON, dSCAVENGER, dCARCASS)) # means specifying the rate of change which has to be in same order as definition of variables
 
})}
```

## Including Plots



```{r}
# to run the model, we need to first define the time steps at which we want to resolve the model.
NYEARS <- 0.15 #number of years of integration
dt <- 0.05 #time step
times <- seq(from =0, to =NYEARS*365, by =dt)
 #using ode from deSolve to solve the model
yini<- States 
for (i in 1:NYEARS) {
  P <- ode (y = yini,
            times =times,
            func = RiverRun,
            parms = Parms) #func means function to integrate and parms means parameters
yini <- P[nrow(P), 2: (length (States)+1)]}

#plotting the state variables and output variables
plot(P, mfrow =c(2, 2), which = 1:4 )
```
#the result shows evolution in time of state variables which an intial condition of 0, salmonn increase and reaches a plato because the source of salmin which is migration reduces at some point and there is increase consumption by bears.
# the carcass a re increasing and not consumed fast enought o reach a plato same as a Bear.
# for scavenger the intial condition can not be mantained byt source and sink, the sink of scavengers are larger than the source the ingestion of carcass at beginign where small but the number increaed as the bear increased and consumption of salmonn increased (due to natural mortality of scavengers)
```{r}
Parms2 <- Parms
Parms2["ingrate"] <- 0.04 # instead of 0.02
yini <- States
for (i in 1: NYEARS){
  P2  <- ode (y = yini,
              times = times,
              func = RiverRun,
              parms = Parms2)
        yini <- P2 [nrow(P2), 2: (length(States)+1)]
}

# Visualizing plot with red color = P2
plot (P, P2, mfrow = c(2, 2), which =1:4)
legend ("topleft", c("ingrate = 0.02", "ingrate = 0.04"), col =1:2, lty = 1:2)
```
#we can observe that salmon doesn't change because the balance of source and sink remains the same because the migration, and the ingestion rate of bear is increasing becasue in the sink the bear ingest increase and the  spawn death reduce but the sink is still the same
#increased flow to carcass and to the scavengers. all this results from a balance between source and sinks
# for the last question you have to multiply by surface